### 2020-06-10 data-raw/development.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

--- 
title: "Issues Developing the new Abalone MSE"
author: "Malcolm Haddon"
date: "`r Sys.Date()`"
font: Times
fontsize: 12pt
output: html_document

---

```{r setup, include=FALSE}
# output: bookdown::gitbook
# bookdown::epub_book: default
# before_body: latex/before_body.tex from output.yml
knitr::opts_chunk$set(size="scriptsize")
knitr::opts_chunk$set(
  cache = FALSE,
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	comment="#",
	dpi = 300,
	fig.align = 'center')

options(knitr.kable.NA = "",
        knitr.table.format = "pandoc")

#wkdir <- "./../../A_Book/book/"

options(show.signif.stars=FALSE,
        stringsAsFactors=FALSE,
        comment="#",
        str = strOptions(strict.width = "cut"))

library(rutilsMH)
library(aMSE)
suppressPackageStartupMessages(library(knitr))
#library(captioner)
library(diagrams)

setpalette("R4")

```

## Main Issues

* How best to define each population in a sequential manner reflective of their relative productivity? Need to keep the option of random allocation of properties. See section.

* The zone is simply to container for all the SAU (there may only be one). The notion is that the variation of population properties within an SAU can differ, but conditioning the populations is a whole subject for discussion in itself.

* Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular zone, there is a need to include the option of having larval dispersal.

* Currently using a control.csv file to identify conditioning files, HCR files, and related. There is a zone file containnig the strucutre of the simulated zone, and a datafile containing the conditioning data.

* When constructing a zone the constant parts are kept in zoneC and the dynamic parts in zoneD. Is the constant part always separate from the global part? Certainly not having to move such large objects around all the time should speed things along.

## Design Principles

* Each run will be conducted in a specified directory, which will always contain at least subdirectories named 'data', 'plots', and 'tables'.

* Each run will be able to generate an html review of all output (see makehtml_utils.R)

* Currently using a spatial hierarchy zone/SAU/population instead of zone/block/population, so this is now more general than only Tasmanian terminology. Stick to this hierarchy throughout the code. Consistency.

* Naming things is always a complex issue, the LML term is still being used, the other States each appear to use different terms: eg. MLL and other three letter acronyms, so stick with LML.

* Once a particular run is constructed, but before the run proceeds, generate a series of plots and tables to characterize the simulated zone.

* All plots should be generated as .png files and stored in the 'plots' directory of each runs directory. In addition, any tables should be saved in the same place as csv files (for matrices) or RData files (for more complex objects) ready to be read in for printing in the html output. Each plot and table will have a unique name that will include the run's name. This will nbe necessary to automate the html output and later for any .Rmd output.

### function: definepops

When generating the properties of each population, currently, the population properties, as defined by the PDfs describing each property, are allocated randomly. With the advent of the spatial data it should be possible to condition the sequence of populations along the coastline in a manner more reflective of the relative productivity of each small area selected to represent a model population.

Random is relatively easy, but both random and specific will be driven by the conditioning. Perhaps best done in relative terms. A detailed examination of hte distribution of ctaches among and between the eastern Block 13 may be illuminating.

Currently the data-file used to condition the definepops function simply lists the mean values of the parameters for all populations per block. It will become necessary to define each populatin spearately to implement sequential definitions. Rather than conditioning entailing continual running until a combination arises that generates about the correct totals, it will be necessary to develop routines to assist searching for suitable levels of productivity for each population. Perhaps there are some simplifiying relationships between Me, G, and AvRec that can aid in this search conditioning.

### larval dispersal

To enable larval dispersal it will be necessary to alter the algorithm whereby each population is dealt with completely before movingon to the next population. Thus, it will be necessary to complete the annual dynamics, except for recruitment, for all populations and only when all are available undertake the recruitment with (or without dispersal). One detailed part will be achieving whatever initial equilibrium conditions are desired (unfished, some given depletion level, etc.).

There is no analytical solution to the initial equilibrium if there is to be larval dispersal.

The dispersal matrix could be initially defined as a matrix and that can be tested to determine whether only the diagonal is filled.

## Highest Level Outline of Abalone MSE

1) Initiate simulated zone - use a ctrl file to define details of data-file, HCR function, and other initial conditions. May involve setting an initial depletion level.

2) conduct assessment of zone status at start of each year (option for partial year?)

3) Using selected HCR set catches for subsequent year(s), and allocate among populations using model fleet dynamics.

4) run the model dynamics for one year

5) If simulation duration not complete return to item 2)

6) If complete then summarize and store results

### Requirements

1) A data-file that contains details for each population

* SAUname    - eg block1, block2
* popnum - population index   1, 2, 3, etc
* MaxDL      - maximum growth increment  eg 32
* sMaxDL     - variation of maxDL  eg  2.5
* L50        - Length at 50% MaxDL  eg 123
* sL50       - variation in L50  eg  5
* L50inc - L95 - L50 =  delta  eg 45
* sL50inc - variation of L50inc  eg 1.25
* SigMax - maximum variation around growth increment eg 4.6
* sSigMax - variation in SigMax  eg 0.1
* Wtb - weight-at-length exponent eg 3.161963
* sWtb - variation of Wtb  eg 0.148
* Wtbtoa - intercept power relationship betweeb Wtb and Wta 962.8098
* sWtbtoa - exponent of power relationship betweeb Wtb and Wta -14.3526
* Me - Natural mortality eg 0.2 maxage = 23, 0.1 maxage = 46 
* sMe - variation in Me  eg 0.003
* AvRec - log of unfished recruitment LnR0  eg 13
* sAvRec -   eg 1.0
* defsteep - Beverton-Holt Steepness  eg 0.6
* sdefsteep         eg 0.02
* L50C - length at 50% emergent  126.422
* sL50C                           1.5
* deltaC - length at increment 95% emergent 95 = 50 + deltaC eg  10
* sdeltaC                                      0.5
* MaxCEpar - maximum cpue tonnes/hour for defining catchability  0.37
* sMaxCEpar                              0.02
* selL50p - L50 of selectivity           0   (deviation from LML)
* selL95p - L95 of selectivity           1.5
* SaMa - maturity logistic a parameter  eg -16
* L50Mat - L50 for maturity b = -a/L50  eg  123.384
* sL50Mat                                   4.0
* larvalD - larval drift                    0.02  = 2%

2) A data-file containing details relating to the whole zone

* numpop - number of populations
* PDFS   - The number of parameters with probability density functions
* minc   - minimum size class  2mm
* cw     - class width         2mm 
* Nclass - number of size classes 105
* nSAU   - number of Spatial Management Units  eg  2
* LML    - current Legal Minimum Length (LML, MLL, MLS, etc) e.g. 140
* randomseed - for repeatability of results - use a different one for each simulation
* Nyrs   - number of years for each simulation   eg 40
* firstYear - first year of simulation eg 1 (hypothetical) or 2014 (conditioned on fishery)
* fixYear - ?  eg 3  used in plotsimulation3
* projLML - the LML to be used in the simulated dynamics

3) A control-file containing details of each run

* label for run
* zone file name
* datafile name
* HCR details file
* replicates
* HCRlabel
* output directory

## The Dynamics Described

### The Initial Equilibrium

The unfished equilibrium is a classical starting point for any model, even if a decision is made to begin any simulations at some lower depletion level. We are using a one-year time-step but the dynamics occurs within each year. The dynamic part of the zone has the following structure:

zoneD - the dynamic part of the zonal structure

* SAU     : num [1:6] 1 1 2 2 2 2
* matureB : num [1:40, 1:6] 782 0 0 0 0 ...
* exploitB: num [1:40, 1:6] 892 0 0 0 0 ...
* catch   : num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* harvestR: num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* cpue    : num [1:40, 1:6] 399 0 0 0 0 ...
* recruit : num [1:40, 1:6] 630563 0 0 0 0 ...
* deplsB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
* depleB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
* catchN  : num [1:105, 1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
* Nt      : num [1:105, 1:40, 1:6] 6.31e+05 2.34e-06 4.31e-05 ...

This represents the dynamics before any simulations have occurred, the unfished equilibrium state. Similarly, the constant part of the zone, made up of a list for each population, has the following structure:

* Me      : num 0.198
* R0      : num 636147
* B0      : num 782
* effB0   : num 782
* ExB0    : num 892
* effExB0 : num 892
* MSY     : num 54.3
* MSYDepl : num 0.304
* bLML    : num 153
* popq    : num 0.000447
* SaM     : num 139
* popdef  : Named num [1:19] 34.45 122.75 164.92 4.61 -16 ....
* LML     : num [1:40] 132 132 132 132 132 132 132 132 132 132 ...
* G       : 'STM' num [1:105, 1:105] 2.12e-13 4.52e-12 8.33e-11 1.27e-09 1.62e-08 ...
* Maturity: num [1:105] 1.42e-07 1.78e-07 2.24e-07 2.83e-07 3.56e-07 ...
* WtL     : num [1:105] 0.000351 0.003438 0.013072 0.033719 0.070319 ...
* Emergent: num [1:105] 1.92e-16 3.44e-16 6.18e-16 1.11e-15 1.99e-15 ...
* Select  : num [1:105, 1:40] 2.87e-127 2.61e-125 2.37e-123 2.16e-121 1.96e-119 ...
* SelWt   : num [1:105, 1:40] 1.01e-130 8.97e-128 3.10e-125 7.28e-123 1.38e-120 ...
* MatWt   : num [1:105] 4.97e-11 6.13e-10 2.93e-09 9.53e-09 2.50e-08 ...
* SAU     : num 1

In the code (_setupzone_) the sequence is:

1. first generate zoneC from the constants and zone1. At that point each population structure contains everything except those things relating to productivity: B0, ExB0, MSY, MSYDepl and popq. makezoneC also updates glb by including the move matrix.
2. running the _makezone_ updates B0, ExB0, and popq in zoneC and generates the initial equilibrium for zoneD. As this is the unfished equilibrium the catch, harvestR, and catchN matrices are empty.
3. The final run of _moderegC_ internally runs _doproductivity_ which allows zoneC's MSY and MSYDepl entries to be updated for each population, as well as producing the product matrix.

In the zoneD matrices, the numbers-at-size are the numbers at the start of the year, similarly, the matureB is the mature biomass at the start of the year. The ExB0 is the exploitable biomass implied by the initial numbers-at-size, which means to calcualte it we need to apply half of natural mortality and growth before multiplyiung by selectivity and weight-at-length.