### 2025-04-09 README.Rmd
Prior to making aMSE public.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aMSE

<!-- badges: start -->
<!-- badges: end -->

This is the codebase for an Abalone Management Strategy Evaluation framework, which we call **aMSE**. It has a novel structure that is based around the spatial ideas of populations (or areas of persistent production or apps, reflecting how they are defined), within spatial management units (SAUs), within a zone:

* zone - highest geographical level. Is simply the totality of the spatial management units.

* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.

* population/app - a population, herein defined as an area of persistent production. These are the active components within the simulation framework. The dynamics of the simulation are based around the populations, although, with positive larval dispersal (a low level is the default) there is some dependency of neighbouring populations on each other.

This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU.

An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant. This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.

Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.

## Installation

You can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:

```{r echo=TRUE, eval=FALSE}
if (!require(devtools)){install.packages("devtools")}

devtools::install_github("https://github.com/haddonm/aMSE")

# similarly for codeutils, hplot, EGHS, and makehtml
# knitr you should get from CRAN
```

The working of the **aMSE** software and how to use it has been documented during the developemt and this is available as a GitBook at:

https://haddonm.github.io/aMSEGuide/

That can ether be read online or a PDF of the whole downloaded (look for the
small Acrobat symbool at top left).

## Example

This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 50 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000). It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.

```{r example, echo=TRUE, eval=FALSE }
# a constant TAC example
starttime <- (Sys.time())
options("show.signif.stars"=FALSE,
        "stringsAsFactors"=FALSE,
        "max.print"=50000,
        "width"=240)
# declare libraries ------------------------------------------------------------
suppressPackageStartupMessages({
  library(aMSE)       # from github.com/haddonm MSE codebase
  library(EGHS)       # from github.com/haddonm example Tas harvest strategy
  library(codeutils)  # from github.com/haddonm a set of code utilities
  library(hplot)      # from github.com/haddonm a set of plotting routines
  library(makehtml)   # from github.com/haddonm used to display results
  library(knitr)      # from CRAN for generating tables
})
dbdir <- getDBdir()   # utility to define the dropbox directory
# Obviously you should modify the rundir and datadir to suit your own setup
prefixdir <- paste0(dbdir,"A_CodeUse/aMSEUse/scenarios/")
postfixdir <- "M15h75"      # a name for your scenario 
rundir <- paste0(prefixdir,postfixdir) # files and results directory
confirmdir(rundir)
controlfile <- "controlM15h75.csv"    # name of the control file
# equilibrium zone -------------------------------------------------------------
# You now need to ensure that there is, at least, a control??.csv, and a 
# saudata??.csv file in the data directory plus some other data .csv files
# depending on how conditioned you want the model to be. Templates for the
# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
# 
# Of course, usually one would use data files, control.csv and a saudata.csv, which
# is listed as the datafile within the control.csv. These must be stored in 
# rundir. Read the Using_aMSE' section in the Documentation .docx file in the 
# documentation. To see examples of code to run the MSE for a real question.

# Define the arguments used in teh harvest strategy
hsargs <- list(mult=0.1,
               wid = 4,
               targqnt = 0.55,
               maxtarg = c(150,150,150,150,150,150,150,150),
               pmwts = c(0.65,0.25,0.1),
               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2),
               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.15,1.2,1.25),
               startCE = 1992,
               endCE = 2020,
               metRunder = 2,
               metRover = 2,
               decrement=1,
               pmwtSwitch = 4,
               stablewts = c(0.4, 0.5, 0.1),
               hcrname="constantrefhcr",
               printmat=NULL)

hsargs <- checkhsargs(hsargs)

deleteyrs <- 0  # none removed in the western zone

# One would need to unhash this to get it to run (obviously)

#out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=constantrefhcr,
#              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
#              getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
#              fleetdyn=NULL,scoreplot=plotfinalscores,
#              plotmultflags=plotmultandflags,interimout="",
#              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
#              matureL=c(40,170),wtatL=c(50,200),mincount=120,
#              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15),
#              nasInterval=1,minsizecomp=c(100,135),uplimH=0.375,incH=0.005,
#              deleteyrs=deleteyrs,selectyr=0)

#makeoutput(out,rundir,postfixdir,controlfile,hsfile="TasHS Package",
#           openfile=TRUE,verbose=FALSE)

# save(out,file=paste0(outdir,postfixdir,".RData"))

```

After running the whole, even if you do not generate the results webpage, you could also try:

```{r }
#str(out$zoneDP,max.level=1)
#str1(out$hcrout)

```

To see the structure of the dynamic object generated by the projections.

See the _aMSEGuide_ for detailed examples.